---
output:
  github_document:
    html_preview: true
---

```{r, include=FALSE, echo=FALSE, warning=FALSE}

library(knitr)
# Set knitr chunk options
opts_chunk$set(include=TRUE, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE, fig.path = "man/figures/")

```

## Animal Density from Camera Data <img src = "man/figures/ABMI-notext.png" height="100" width="100" align="right" />

> Data exploration, methods description, and functions from {abmi.camera.extras}

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/mabecker89/abmi.camera.extras.svg?branch=master)](https://travis-ci.org/mabecker89/abmi.camera.extras)
<!-- badges: end -->

This package and content provide information related to how the Alberta Biodiversity Monitoring Institute (ABMI) estimates the density of animals using data collected from remote cameras. 

### Site Contents

#### ABMI Camera Data

  + [Image summaries](https://mabecker89.github.io/abmi.camera.extras/)
  + [Species detections](https://mabecker89.github.io/abmi.camera.extras/)

#### Methods

  + [Overview](link)
  + [Probabilistic gaps](https://mabecker89.github.io/abmi.camera.extras/)
  + [Deployment timeframes](https://mabecker89.github.io/abmi.camera.extras/)
  + [Time in the camera field of view](https://mabecker89.github.io/abmi.camera.extras/)
  + [Effective detection distance](https://mabecker89.github.io/abmi.camera.extras/)
  + [Density at individual deployments](https://mabecker89.github.io/abmi.camera.extras/)
  + [Lure adjustments](https://mabecker89.github.io/abmi.camera.extras/)
  + [Density in an area of interest (AOI)](https://mabecker89.github.io/abmi.camera.extras/)
  + [Important assumptions](https://mabecker89.github.io/abmi.camera.extras/)

#### Functions

## Installation

```{r, include=TRUE, eval=FALSE}

# Install the latest version from Github:
# install.packages("remotes")
remotes::install_github("mabecker89/abmi.camera.extras")

```

## License 

The estimates, predictions, and related documentation are © ABMI (2013-2019) under a [CC BY-SA 4.0 license](https://creativecommons.org/licenses/by-sa/4.0/).

This R package is licensed under [MIT license](https://github.com/mabecker89/abmi.camera.extras/blob/master/LICENSE) © 2020 Marcus Becker, David Huggard, and the ABMI.

## Issues

To report bugs, request additional features, or get help using the package, please file an [issue](https://github.com/mabecker89/abmi.camera.extras/issues). You can also email Marcus Becker mabecker@ualberta.ca. 






